Topic modeling on EU and the USA research projects

### 2017-06-09 thesis_report/draft_thesis.tex
Update: some probability concepts

\title{Topic modeling bla bla}

@@

\subsection{The Multinomial Distribution}

@@

@@

This statistical model reflects the intuition that documents exhibit multiple
topics. Each document exhibits the topics with different proportion (step 1);
each word in each document is drawn from one of the topics (step 2b), where the
selected topic is chosen from the per-document distribution over topics (step
2a). All the documents in the corpus share the same set of topics, but each
document exhibits those topics with different proportion - that’s the
distinguishing characteristic of LDA. A more detailed explanation of the model
follows in section 2.3.3.2.

\paragraph{Dirichlet distribution}

>>>>

\title{Topic modeling on EU and the USA research projects}
\author{Maria Iliadi}

@@

\subsection{Some probability concepts}

\subparagraph{Prior and posterior distributions}

When introducing the Latent Dirichlet Allocation in section 2.x.x, we will be
using the terms prior and posterior probabilities and likelihood function:

\begin{itemize}
\item The prior probability distribution of and uncertain quantity is the 
probability distribution that express the value of this quantity before 
some evidence is taken into account.
\item The posterior probability is the conditional probability after some 
evidence is taken into account.
\item The likelihood function of a set of parameters given some outcome 
is equal to the probability of observing said outcome given those 
parameter values.
\end{itemize}

If the prior and the posterior distributions are in the same distribution
family, then the prior and posterior are called conjugate distributions and the
prior is called a conjugate prior for the likelihood funtion.

Given a prior $p(\theta)$ and observations x with the likelihood 
$p(x \mid \theta)$, the posterior is defined as:

\begin{equation}
p(\theta \mid x) = \frac{p(x \mid \theta) p(\theta)}{p(x)}
\end{equation}

\subparagraph{Jensen Inequality}

In the context of probabilities, the Jensen Inequality is stated as follows: 
If X is a random variable and f a convex function then:

\begin{equation}
f(\mathbb{E}[X]) \leq \mathbb{E}[f(X)]
\end{equation}

And, if f is a concave function the inequality turns:

\begin{equation}
f(\mathbb{E}[X]) \geq \mathbb{E}[f(X)]
\end{equation}

\subparagraph{Kullback-Leibler (KL) divergence}

When formulating the optimization problem in Latent Dirichlet Allocation we 
will use the KL divergence to express the optimization problem. The KL divergence 
is a measure of the difference between two probability distributions $P$ and $Q$. 
It is important to note
that the KL divergence is not symmetrical in $P$ and $Q$ so it’s not a metric.

The KL divergence can be interpreted as a way to express the amount of 
information lost when $Q$ is used to approximate $P$.

Given two continuous distributions $P$ and $Q$ the KL divergence is defined as follows:

\begin{equation}
D_{KL}(P \mid \mid Q) = \int_{-\infty}^{+\infty} p(x) log\frac{p(x)}{q(x)}dx
\end{equation}

This can also be expressed using the definition of expectation (and expressing the 
log of a division as difference of log) as:

\begin{equation}
D_{KL}(P \mid \mid Q) = \mathbb{E}_{p}[log p(x)] - \mathbb{E}_{p}[log q(x)]
\end{equation}

This is the expression we will use later on instead of the definition. We present 
now the two probability distributions we will be using when we define the different 
models in this work: the \textbf{Multinomial} and the \textbf{Dirichlet} distributions.

\subparagraph{Multinomial Distribution}

@@

\subparagraph{Dirichlet distribution}

@@

This statistical model reflects the intuition that documents exhibit multiple
topics. Each document exhibits the topics with different proportion (step 1);
each word in each document is drawn from one of the topics (step 2b), where the
selected topic is chosen from the per-document distribution over topics (step
2a). All the documents in the corpus share the same set of topics, but each
document exhibits those topics with different proportion - that’s the
distinguishing characteristic of LDA. A more detailed explanation of the model
follows in section 2.3.3.1.
